---
title: "Part 2: Chi Square with lizard data"
author: "Jessica Madden"
date: "December 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species:
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST" (east tarbus site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```

```{r}
#janitor packaage has funcitons that start to create contingency tables which are not tidy but better for communication and chi sq

tail_counts <- utst %>% 
  tabyl(site, tail) 

tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
```

## Convert to a contingency table

Use `column_to_rowname()`.

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square

null hypothesis is that proportions of whole to broken tails would be equal between all sites (independence)

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

p<0.05 is sufficient evidence to reject the null hypothesis

there is a significant association/effect of between site and tail breakage; site and lizard tail breakge arenot indepenent

There is a significant association (i.e. non-independence) betwen site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).
